---
title: "Matching signature to phenotype"
author: "Sara Gosline"
date: "11/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

To compare a signature for a particular patient population phenotype, we use the patient distance metrics defined in the previous section to identify specific networks that correlate with a phenotype of interest. 

## Signature selection

We can now use the network signatures to determine which signatures best correlate with a specific phenotype using a basic linear modeling approach. 

We can start with identifying which signatures best predict cancer type. 

```{r signature selection}

```

## Incorporating outside data

Cancer is by definition a genetic disease, yet genetic mutations are often difficult to detect. By using these protein signatures we can evaluate broader phenotypes.


```{r genetic mtuation data}

##read in data here

```

## Genetic mutation signatures span cancer types

We look for specific signatures and can stratify patients?

## Mutation signatures also recapitulate known biology

We can also recapitulate known biology

